---
title: "Quick Data Visualization"
author: "Renata"
date: "2024-07-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##Set-up

```{r read in data}

ChicagoData <- read.csv("./ChicagoSegmentData.csv")
DurhamData <- read.csv("./DurhamSegmentData.csv")


```



###Look at Census data

Change Median Household Income
Median Household Income

Change Percent White non-HL
Percent White non-HL

Avg Percent Renter

Avg Percent Recent Movers

SVI

Construction

% commercial

% residential

```{r Census data}

##case study areas - Chicago
Cgr1 <- c(57, 16)
Cgr2 <- c(244, 49)
ChiCaseStudyData1 <- ChicagoData[ChicagoData$Segment %in% Cgr1,] #blues
ChiCaseStudyData2 <- ChicagoData[ChicagoData$Segment %in% Cgr2,] #reds

#####case study areas - Durham
Dgr1 <- c("M22", "K18") #greens
Dgr2 <- c("O21", "M19") #purples
Dgr3 <- c("R21", "R24") #yellows

DurCaseStudyData1 <- DurhamData[DurhamData$Segment %in% Dgr1,]
DurCaseStudyData2 <- DurhamData[DurhamData$Segment %in% Dgr2,]
DurCaseStudyData3 <- DurhamData[DurhamData$Segment %in% Dgr3,]

#############Median Household Income
hist(ChicagoData$MedHouseInc21, breaks = 15, main="Chicago Median Household Income 2021")
abline(v=ChiCaseStudyData1$MedHouseInc21[1], col="cyan", lwd=3)
abline(v=ChiCaseStudyData1$MedHouseInc21[2], col="blue", lwd=3)
abline(v=ChiCaseStudyData2$MedHouseInc21[1], col="coral", lwd=3) #244
abline(v=ChiCaseStudyData2$MedHouseInc21[2], col="red", lwd=3) #49

hist(ChicagoData$Change_MedHHI, breaks = 10, main="Chicago Change Median Household Income")
abline(v=ChiCaseStudyData1$Change_MedHHI[1], col="cyan", lwd=3)
abline(v=ChiCaseStudyData1$Change_MedHHI[2], col="blue", lwd=3)
abline(v=ChiCaseStudyData2$Change_MedHHI[1], col="coral", lwd=3)
abline(v=ChiCaseStudyData2$Change_MedHHI[2], col="red", lwd=3)

hist(DurhamData$MedHouseInc21, breaks = 20, main="Durham Median Household Income 2021")
abline(v=DurCaseStudyData1$MedHouseInc21[1], col="springgreen", lwd=3)
abline(v=DurCaseStudyData1$MedHouseInc21[2], col="green4", lwd=3)

hist(DurhamData$Change_MedHHI, breaks = 10, main="Durham Change Median Household Income")


#############Percent White non-HL
hist(ChicagoData$PercWhiteNonHL21, breaks = 15, main="Chicago Percent White 2021")
abline(v=ChiCaseStudyData1$PercWhiteNonHL21[1], col="cyan", lwd=3)
abline(v=ChiCaseStudyData1$PercWhiteNonHL21[2], col="blue", lwd=3)
abline(v=ChiCaseStudyData2$PercWhiteNonHL21[1], col="coral", lwd=3)
abline(v=ChiCaseStudyData2$PercWhiteNonHL21[2], col="red", lwd=3)
hist(ChicagoData$Change_PWnHL, breaks = 10, main="Chicago Change Percent White")
abline(v=ChiCaseStudyData1$Change_PWnHL[1], col="cyan", lwd=3)
abline(v=ChiCaseStudyData1$Change_PWnHL[2], col="blue", lwd=3)
abline(v=ChiCaseStudyData2$Change_PWnHL[1], col="coral", lwd=3)
abline(v=ChiCaseStudyData2$Change_PWnHL[2], col="red", lwd=3)


hist(DurhamData$PercWhiteNonHL21, breaks = 20, main="Durham Percent White")
hist(DurhamData$Change_PWnHL, breaks = 10, main="Durham Change Percent White")


###################Average Percent Renter
hist(ChicagoData$Avg_PRent, breaks = 10, main="Chicago Avg Percent Renter")
abline(v=ChiCaseStudyData1$Avg_PRent[1], col="cyan", lwd=3)
abline(v=ChiCaseStudyData1$Avg_PRent[2], col="blue", lwd=3)
abline(v=ChiCaseStudyData2$Avg_PRent[1], col="coral", lwd=3)
abline(v=ChiCaseStudyData2$Avg_PRent[2], col="red", lwd=3)

hist(DurhamData$Avg_PRent, breaks = 10, main="Durham Avg Percent Renter")


####################Average Recent Movers
hist(ChicagoData$Avg_Rmov, breaks = 10, main="Chicago Avg Percent Recent Movers")
abline(v=ChiCaseStudyData1$Avg_Rmov[1], col="cyan", lwd=3)
abline(v=ChiCaseStudyData1$Avg_Rmov[2], col="blue", lwd=3)
abline(v=ChiCaseStudyData2$Avg_Rmov[1], col="coral", lwd=3)
abline(v=ChiCaseStudyData2$Avg_Rmov[2], col="red", lwd=3)

hist(DurhamData$Avg_Rmov, breaks = 10, main="Durham Avg Percent Recent Movers")


######################SVI
hist(ChicagoData$SVI, breaks = 10, main="Chicago SVI")
abline(v=ChiCaseStudyData1$SVI[1], col="cyan", lwd=3)
abline(v=ChiCaseStudyData1$SVI[2], col="blue", lwd=3)
abline(v=ChiCaseStudyData2$SVI[1], col="coral", lwd=3)
abline(v=ChiCaseStudyData2$SVI[2], col="red", lwd=3)

hist(DurhamData$SVI, breaks = 10, main="Durham SVI")


####################Construction
hist(ChicagoData$AllConst_since11, breaks = 10, main="Chicago All Construction")
abline(v=ChiCaseStudyData1$AllConst_since11[1], col="cyan", lwd=3)
abline(v=ChiCaseStudyData1$AllConst_since11[2], col="blue", lwd=3)
abline(v=ChiCaseStudyData2$AllConst_since11[1], col="coral", lwd=3)
abline(v=ChiCaseStudyData2$AllConst_since11[2], col="red", lwd=3)

hist(DurhamData$NumConst_since11, breaks = 10, main="Durham Construction since 2011")
hist(DurhamData$NumReno_since11, breaks = 10, main="Durham Renovation since 2011")

```

##Notes

Chicago:
* blues - generally more similar in income and percent white in 2021, with different trajectories to get there and separation in avg renters and avg recent movers
  - light blue has had more change in income and percent white but has lower rentership and recent movers
* reds - have a bit more of a gap in 2021 in income, but the change is more similar, similar both percent white and change. Pretty big difference in average recent movers, and the SVI is super similar

Makes me consider focusing on the reds, being able to look at the contrast with recent movers and  construction?


Durham
* only focusing on the segments with 10+ trees
* get rid of Q21 because it is almost entirely commercialand O19 and P21 because they are mostly ot residential